Cognitive and ecological aspects determine cleaner performance

### 2021-11-10 manuscript_1.0.Rmd
Erased old simulations which were no longer needed. Added Redouan intro to the manuscript_. New Chains were parameters can be negative.

---
title: "Cognitive and ecological aspects determine cleaner performance"
author:
- Andrés E. Quiñones^[Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland - andresqp@gmail.com]
- Zegni Triki^[Department of Zoology, Stockholm University, Stockholm, Sweden]
- Redouan Bshary^[Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland]
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  word_document: default
  pdf_document: default
fig_caption: yes
fontsize: 12pt
header-includes:
- \usepackage{float}
- \usepackage{setspace}\doublespacing
- \linespread{2}
- \usepackage{lineno}
- \linenumbers
- \newcommand{\beginsupplement}{
    \setcounter{table}{0}  
    \renewcommand{\thetable}{S\arabic{table}} 
    \setcounter{figure}{0} 
    \renewcommand{\thefigure}{S\arabic{figure}}
keep_tex: yes
keywords: My keywords
linespread: 2
bibliography: Cleaner_learning.bib
style: C:\Users\andre\Zotero\styles\behavioral-ecology-and-sociobiology.csl
spacing: double
abstract: Insert a nice abstract here.
---

@@

Cooperation among individuals of the same or different species is ubiquitous in nature. 
Among individuals of the same species it is the basis for the evolution of  higher 
forms of biological organization [@maynardsmith_Major_1998]. 
Among individuals of different species it underlays the biological mechanism for 
fundamental ecological interactions such as pollination, nitrogen fixation or 
parasite removal [@begon_Ecology_2006]. However, when cooperation is costly, 
it is not straightforward to understand why individuals would be willing to pay that cost [@lehmann_Evolution_2006]. Furthermore, cooperative interactions might open the way for 
one partner to exploit the other, what prevents this type of exploitation? [@herre_Evolution_1999]
One reason for cooperating is if the benefits obtained from a partner´s 
cooperative action are dependent on one's own cooperative act. 
This dependency can arise, for instance, when cooperative partners
choose who they interact with depending on the level of cooperation they offer. 
In line with this, cooperative interactions can be viewed as a market-like situation 
where partner choice mediates the level of exchange between traders [@hammerstein_Biological_2016; @leimar_Cooperation_2010; @noe_Biological_1994a]. Empirical evidence shows
that several iconic examples of cooperation between individuals of different species 
are indeed mediated by partner choice [@hammerstein_Biological_2016]. Biological market theory highlights 
that changes in the supply and demand of goods and services determines the value of 
those commodities, and therefore, can influence partner's actions in the cooperative 
exchange. For instance, in the marine cleaning mutualism, where cleaner fish 
_Labroides dimidiatus_ (cleaners) removes ectoparasites from other coral reef fishes (clients) [@cote_Evolution_2000],
some client species are able to choose among different cleaners, and because of this
obtain better service [@grutter_Cleaner_2003]. Better services means for clients to have priority when 
more than one client is available [@bshary_Choosy_2002], and less cheating (when cleaners eat live tissue
instead of parasites) [@grutter_Cleaner_2003]. However, recent evidence shows that when clients experience
ecological conditions with low cleaner abundance, they are less likely to use partner
choice to leverage better service [@triki_Biological_2019]. On the other side, 
cleaners in those markets with favourable conditions, stop giving priority to choosy clients [@triki_Biological_2019; @triki_Decrease_2018]. Thus, it seems that cleaners, as well as clients,
adjust their strategy to the prevalent market conditions they find in their environment. 
This adjustment cannot be due to local genetic adaptation, given that cleaners are open water spawners 
and the environmental changes in different sites have occurred in the last decade [@triki_Biological_2019; @triki_Decrease_2018]. Alternatively, a recent cognitive model proposes that differences in 
cleaners preferences can arise trough the interaction of cleaner's cognitive processes 
with the environmental conditions [@quinones_Reinforcement_2019]. Under this view,
cleaners ontogenetically develop their preferences for different clients through learning,
and some environmental conditions facilitate learning. Thus, the adaptive adjustment 
seen in cleaners might be due to their learning mechanisms. Despite the general fit of the 
model explanation to the available data, no systematic parameter fit has been performed to 
evaluate which combination of model parameters better fit the data.

@@

```{r post, fig.cap="Posterior distributions for parameter values $\\gamma$, $\\eta$ and the scalling constant. We show the kerne density estimates, and bellow the mean (black point) and the 65\\% and 95\\%  highest posterior density interval for the three paramteres. On the left (panels a, c and c) posterior distributions for a model with all three parameters; while on the right (b and d) a model without negative reward, ",out.width='85%'}
knitr::include_graphics(here("post_both_gamma.png"))

@@

The main reason for negative and future reward to give different predictions is the way that cleaner's relative abundance influences the preference for the visitor clients. Visitor leaving probability has a similar positive effect on the probability of choosing the visitor clients on all three models 
(Fig. \ref{fig:pred}, \ref{fig:nogamma}). In contrast, cleaner's relative abundance has a different effect in the model with only future reward, compared to the other two. In the model with only future reward 
only intermediate cleaner abundance triggers a preference for the visitor clients 
(Fig. \ref{fig:pred} c). In models with negative reward, both intermediate and low cleaner abundance
trigger a preference for the visitor. (Fig. \ref{fig:pred} a, \ref{fig:nogamma})

@@

<div id="refs"></div>

@@

\beginsupplement

```{r nogamma, fig.cap="Observed and predicted probability of choosing a visitor for a model without future rewards. Left-hand side panel: color contour shows the prediction of the learning model using the mode of the posterior distributions of parameters recovered by the statistical analysis. Dots show the frequency of visitor choices for the 12 locations, as well as the corresponding relative cleaner abundance (x axis) and frequency of visitors leaving the station (y axis). Right-hand side panel: scatter plot of the observed and predicted probabilities of choosing a visitor for 12 locations. Line correspond to a perfect match between observed and predicted probabilities.",out.width='90%'}
fileNogamma<-grep("contour_gamma_",list.files(here("Simulations","ABCclean_nRew_sca_"),
                                              full.names = TRUE),value = TRUE)
knitr::include_graphics(fileNogamma)

>>>>

---
title: "Cognitive and ecological aspects determine cleaner performance"
author:
- Andrés E. Quiñones^[Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland - andresqp@gmail.com]
- Zegni Triki^[Department of Zoology, Stockholm University, Stockholm, Sweden]
- Redouan Bshary^[Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland]
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  pdf_document: default
fig_caption: yes
fontsize: 12pt
header-includes:
- \usepackage{float}
- \usepackage{setspace}\doublespacing
- \linespread{2}
- \usepackage{lineno}
- \linenumbers
- \newcommand{\beginsupplement}{
    \setcounter{table}{0}  
    \renewcommand{\thetable}{S\arabic{table}} 
    \setcounter{figure}{0} 
    \renewcommand{\thefigure}{S\arabic{figure}}}
keep_tex: yes
keywords: My keywords
linespread: 2
bibliography: Cleaner_learning.bib
style: M:\Zotero\styles\behavioral-ecology-and-sociobiology.csl
spacing: double
abstract: Insert a nice abstract here.
editor_options: 
  markdown: 
    wrap: 72
---

@@

Biological market theory emphasizes the importance of partner choice in
explaining how supply-to-demand ratios translate into payoff asymmetries
between partners [@noe_Biological_1994a; @noe_Biological_1995a]. One
example is the marine cleaning mutualism involving the cleaner
fish *Labroides dimidiatus* and 'client' fish. Clients offer themselves as
food patches for cleaners to remove ectoparasites, which gives the
cleaners calories and the clients improved health
[@waldie_LongTerm_2011; @ros_Does_2011,@triki_Effects_2016]. As
cleaners have about 2000 interactions per day
[@grutter_Relationship_1995], they regularly face the situation that
two clients seek cleaning simultaneously. In such conditions, it is
important to distinguish between 'resident' and 'visitor' client species
[@bshary_Cleaner_2001a]. The residents have small home ranges and hence
access to only the local cleaner, while the visitors have large home
ranges and hence access to two or more cleaners. Thus, visitors have the
option to switch to another cleaner if being made to wait, while
residents must wait for inspection. Indeed, visitors have been observed
to use their partner choice options in that way [@bshary_Choosy_2002],
which may explain why cleaners give visitors service priority in a field
study in the Red Sea [@bshary_Cleaner_2001a]. Furthermore, early
laboratory experiments showed that leaving causes service priority
[@bshary_Asymmetric_2002]. The experiments consisted of a simultaneous
two-choice task, where cleaners had to decide which of two Plexiglas
plates to approach first and eat mashed prawn of it. One plate was made
to behave like a resident, i.e. it would remain until the cleaner had
eaten off it. The other plate was made to behave like a visitor, i.e. it
would be removed if not inspected first. Cleaners learned to prefer the
visitor plate and hence obtained the double amount of food
[@bshary_Asymmetric_2002]. Adult cleaner fish outperform various
primates as well as rats and pigeons in this original version of the
biological market task [@salwiczek_Adult_2012; @zentall_Early_2017];
only African grey parrots solve this task as well [@pepperberg_Can_2014].

Recent research has refined our view on partner choice and priority of
access in the *L. dimidiatus* cleaning mutualism, and as a consequence
performance in the biological market task. First, details of the task
presentation may matter; i.e. what relevant information is encoded in
the design of the two plates (colour versus size; @wismer_Cuebased_2019), or
whether the information is in the colour of the food
[@pretot_Comparative_2021; @pretot_Comparing_2016], and whether the
food is visible or hidden/released only when the subjects made choices
on a touch screen [@pretot_Factors_2016]. Second, it appears that cleaner and
client densities in nature affect how cleaners perform in the original
market task: when densities are low, cleaners are more likely to fail to
complete the task [@triki_Biological_2019; @triki_Decrease_2018;
@wismer_Variation_2014]. The apparent explanation for this failure is that visitor
clients are less likely to exert their choice options (and hence wait
for inspection if made to wait) when cleaner fish densities are low
[@triki_Brain_2020]. This change in behavior may be due to the fact that
when cleaner densities are low, the cost of switching in terms of
distance to swim and time invested increases. If visitors are not
choosy, cleaners should not give them priority anymore. Interestingly,
while cleaners from high-density areas have larger forebrains harbouring
more cells than cleaners from low-density areas [@triki_Brain_2020], on a
local scale those individuals that perform in the market task in a way
that corresponds to the locally optimal strategy (solving/failing the
task when living in high/low density) have relatively larger forebrains
[@triki_Brain_2020]. Taken together, recent empirical results clearly show
that we need a mechanistic approach to understand variation in intra-
and interspecific performance in the market task. One approach to
identify candidate mechanisms underlying performance in a cognitive task
is to develop proximate learning models. By varying the cognitive tool
kit, one can assess which minimal kit is necessary to solve the task at
hand. This approach has been used to analyse the market task
[@prat_Role_2021; @quinones_Reinforcement_2019] and spatial learning in bees [@dubois_Model_2021]. Learning models showed that basic reinforcement 
learning does not suffice to solve the task, this is particularly so when model assume the more complex natural situation in which cleaners
not only face resident-visitor pairs but also visitor-visitor and
resident-resident pairs, as well only a resident or only a visitor. To
be able to give visitors priority over residents, cleaners first need to
be able to assess a client's value separately for the three possible
forms of encounter (alone, paired with a fish with the same strategic
option, paired with a fish with the alternative strategic option)
[@quinones_Reinforcement_2019]. In addition, cleaners need to account for the future
consequences of current decisions. In the model, this could be achieved
in two non-mutually exclusive ways [@quinones_Reinforcement_2019]:
through low temporal discounting of future effects
and/or through perceiving a visitor client leaving as psychological
punishment (i.e. as a negative reinforcer). Given that a negative
reinforcer is part of the ubiquitous learning mechanism termed operant
conditioning (Thorndike xxxx), this explanation can be considered the
simpler one. Furthermore, accounting for the future effects of an action is 
considered an advanced cognitive ability [@Suddenford_Evolution_2007]. 
Here, we fitted our model to the empirical data on fish
densities and cleaner fish performance in the market task with the aim
to determine the likely mechanism used by cleaners to incorporate future
consequences of current decisions.

Cooperation among individuals of the same or different species is
ubiquitous in nature. Among individuals of the same species it is the
basis for the evolution of higher forms of biological organization
[@maynardsmith_Major_1998]. Among individuals of different species it
underlays the biological mechanism for fundamental ecological
interactions such as pollination, nitrogen fixation or parasite removal
[@begon_Ecology_2006]. However, when cooperation is costly, it is not
straightforward to understand why individuals would be willing to pay
that cost [@lehmann_Evolution_2006]. Furthermore, cooperative
interactions might open the way for one partner to exploit the other,
what prevents this type of exploitation? [@herre_Evolution_1999] One
reason for cooperating is if the benefits obtained from a partner´s
cooperative action are dependent on one's own cooperative act. This
dependency can arise, for instance, when cooperative partners choose who
they interact with depending on the level of cooperation they offer. In
line with this, cooperative interactions can be viewed as a market-like
situation where partner choice mediates the level of exchange between
traders [@hammerstein_Biological_2016; @leimar_Cooperation_2010;
@noe_Biological_1994a]. Empirical evidence shows that several iconic
examples of cooperation between individuals of different species are
indeed mediated by partner choice [@hammerstein_Biological_2016].
Biological market theory highlights that changes in the supply and
demand of goods and services determines the value of those commodities,
and therefore, can influence partner's actions in the cooperative
exchange. For instance, in the marine cleaning mutualism, where cleaner
fish *Labroides dimidiatus* (cleaners) removes ectoparasites from other
coral reef fishes (clients) [@cote_Evolution_2000], some client
species are able to choose among different cleaners, and because of this
obtain better service [@grutter_Cleaner_2003]. Better services means
for clients to have priority when more than one client is available
[@bshary_Choosy_2002], and less cheating (when cleaners eat live
tissue instead of parasites) [@grutter_Cleaner_2003]. However, recent
evidence shows that when clients experience ecological conditions with
low cleaner abundance, they are less likely to use partner choice to
leverage better service [@triki_Biological_2019]. On the other side,
cleaners in those markets with favourable conditions, stop giving
priority to choosy clients [@triki_Biological_2019;
@triki_Decrease_2018]. Thus, it seems that cleaners, as well as
clients, adjust their strategy to the prevalent market conditions they
find in their environment. This adjustment cannot be due to local
genetic adaptation, given that cleaners are open water spawners and the
environmental changes in different sites have occurred in the last
decade [@triki_Biological_2019; @triki_Decrease_2018]. Alternatively, a
recent cognitive model proposes that differences in cleaners preferences
can arise trough the interaction of cleaner's cognitive processes with
the environmental conditions [@quinones_Reinforcement_2019]. Under this
view, cleaners ontogenetically develop their preferences for different
clients through learning, and some environmental conditions facilitate
learning. Thus, the adaptive adjustment seen in cleaners might be due to
their learning mechanisms. Despite the general fit of the model
explanation to the available data, no systematic parameter fit has been
performed to evaluate which combination of model parameters better fit
the data.

@@

```{r post, fig.cap="Posterior distributions for parameter values $\\gamma$, $\\eta$ and the scalling constant. We show the kerne density estimates, and bellow the mode (black point) and the 65\\% (light blue) and 95\\% (grey)  highest posterior density interval for the three paramteres. On the left (panels a, c and c) posterior distributions for a model with all three parameters; while on the right (b and d) a model without negative reward, ",out.width='85%'}
knitr::include_graphics(here("post_both_gamma.png"))

@@

The main reason for negative and future reward to give different
predictions is the way that cleaner's relative abundance influences the
preference for the visitor clients. Visitor leaving probability has a
similar positive effect on the probability of choosing the visitor
clients on all three models (Fig. \ref{fig:pred}, \ref{fig:nogamma}). In
contrast, cleaner's relative abundance has a different effect in the
model with only future reward, compared to the other two. In the model
with only future reward only intermediate cleaner abundance triggers a
preference for the visitor clients (Fig. \ref{fig:pred} c). In models
with negative reward, both intermediate and low cleaner abundance
trigger a preference for the visitor. (Fig. \ref{fig:pred} a,
\ref{fig:nogamma}).

@@

::: {#refs}
:::

@@

```{=tex}
\beginsupplement
```
```{r nogamma, fig.cap="Observed and predicted probability of choosing a visitor for a model without future rewards. Left-hand side panel: color contour shows the prediction of the learning model using the mode of the posterior distributions of parameters recovered by the statistical analysis. Dots show the frequency of visitor choices for the 12 locations, as well as the corresponding relative cleaner abundance (x axis) and frequency of visitors leaving the station (y axis). Right-hand side panel: scatter plot of the observed and predicted probabilities of choosing a visitor for 12 locations. Line correspond to a perfect match between observed and predicted probabilities.",out.width='90%'}
fileNogamma<-grep("contour_gamma_",list.files(here("Simulations","ABCclean_nRew_sca_"),
                                              full.names = TRUE),value = TRUE)
knitr::include_graphics(fileNogamma)
